---
title: "Aaron Capstone Project"
author: "Aaron  Nwafor"
date: "2025-10-06"
output: html_document
---

##Installing packages
```{r setup, include=FALSE}
install.packages(c("tidyverse","haven","survey","srvyr","tableone","gtsummary","janitor","gt"))

```



##loading packages
```{r}
library(tidyverse)
library(haven)    
library(survey)     
library(srvyr)      
library(tableone)   
library(gtsummary)
library(janitor)
library(gt)

```

##Reading my file path
```{r}
Capstone <- haven::read_sav("C:/Users/Aaron/Desktop/CapstonefirearmBRFSS - Copy - Copy.sav")
```



```{r}
library(dplyr)
var_SEXVAR <- "_SEX"

Capstone
```

##Renaming
```{R}
df <- Capstone %>%
  select(all_of(c("MARITAL", "INCOME3", "EDUCA", "FIREARM5", "SEXVAR", "@_RACEGR3", "@_AGE_G", "@_URBSTAT"))) %>%
  rename(
    FIREARMS_Ownership = FIREARM5,
    EDUCATIONAL_STATUS = EDUCA,
    INCOME_LEVELS  = INCOME3,
    URBANICITY   = '@_URBSTAT',
    AGE_GROUPS = '@_AGE_G',
    sEX = SEXVAR,
    RACE = '@_RACEGR3',
    MARITAL_STATUS = MARITAL 
  )

```



##Recode Firearm
```{r}
Capstone_ex <- Capstone%>%
mutate(
    FIREARMS_Ownership = case_when(
      FIREARM5 == 1 ~ "Yes",
      FIREARM5 == 2 ~ "No",
      FIREARM5 %in% c(7, 9) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    
     Educational_level = case_when(
      EDUCA %in% c(1,2,3) ~ "Elementary/Less than High School graduate",  
      EDUCA %in% c(4) ~ "High school graduate/GED",
      EDUCA %in% c(5) ~ "Some college or Technical school",
      EDUCA %in% c(6) ~ "College graduate",
      EDUCA %in% c(9) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    Income_level = case_when(
      INCOME3 %in% c(1,2,3,4,5) ~ "Low(<35K)",
      INCOME3 %in% c(6,7) ~ "Medium ($35k– < $75k)",
      INCOME3%in% c(8,9,10,11) ~ "High(>$75K)",
      INCOME3 %in% c(77) ~ NA_character_,
      TRUE ~ NA_character_
    ),
     URBANICITY = case_when(
      `@_URBSTAT` %in% c(1) ~ "Urban counties",
      `@_URBSTAT` %in% c(2) ~ "Rural counties",
      TRUE ~ NA_character_
     ),
     AGE_YRS = case_when(
      `@_AGE_G` %in% c(1) ~ "18 - 24",  
      `@_AGE_G` %in% c(2) ~ "25 - 34",
      `@_AGE_G` %in% c(3) ~ "35 - 44",
      `@_AGE_G`%in%  c(4) ~ "45 - 54",
      `@_AGE_G`%in%  c(5) ~ "55 - 64",
      `@_AGE_G`%in% c(6) ~  "65 - Older",
      TRUE ~ NA_character_
    ),
    SEX_OF_RESPONDENT = case_when(
      `SEXVAR` %in% c(1) ~ "Male",
      `SEXVAR` %in% c(2) ~ "Female",
      TRUE ~ NA_character_
    ),
    Race_ethnicity = case_when(
      `@_RACEGR3` %in% c(1) ~ "Non Hispanic White",
      `@_RACEGR3` %in% c(2) ~ "Black only",
      `@_RACEGR3` %in% c(3,4) ~ "Multiracial and other race(Non Hispanic)",
      `@_RACEGR3` %in% c(5) ~ "Hispanics",
      '@_RACEGR3' %in% c(9) ~ NA_character_,
      TRUE ~ NA_character_
   ),
    Marital_Status = case_when(
      MARITAL %in% c(1) ~ "Married",
      MARITAL %in% c(2,3,4) ~ "Previously Married(Divorced/Widowed/Seperated)",
      MARITAL %in% c(5) ~ "Unmarried/Single",
      MARITAL %in% c(6) ~ "Unmarred couple",
      MARITAL %in% c(9) ~ NA_character_,
      TRUE ~ NA_character_
    ))

Capstone_ex
    
```



##Checking for missing for Missingness and sample size
```{r}
Capstone_ex1<-Capstone_ex %>% summarise(
  unweighted_n = n(),
  n_FIREARMS_Ownership_missing = sum(is.na(FIREARMS_Ownership)),
  n_Educational_level_missing = sum(is.na(Educational_level)),
  n_Income_level_missing = sum(is.na(Income_level)),
  n_URBANICITY_missing = sum(is.na(URBANICITY)),
  n_AGE_YRS_missing = sum(is.na(AGE_YRS)),
  n_SEX_OF_RESPONDENT = sum(is.na(SEX_OF_RESPONDENT)),
  n_Race_ethnicity_missing = sum(is.na(Race_ethnicity)),
  n_Marital_Status_missing =sum(is.na(Marital_Status))
  
)


```

##Checking all missingness
```{r}
colSums(is.na(Capstone_ex))
```

##Isolate my variables from the Dataset
```{r}
Capstone_1 <- Capstone_ex %>% 
  select(all_of(c("FIREARMS_Ownership","Educational_level","Income_level","URBANICITY","AGE_YRS","SEX_OF_RESPONDENT","Race_ethnicity","Marital_Status")))

```



##Checking for missing for Missingness and sample size
```{r}
Capstone_1mi<-Capstone_1 %>% summarise(
  unweighted_n = n(),
  n_FIREARMS_Ownership_missing = sum(is.na(FIREARMS_Ownership)),
  n_Educational_level_missing = sum(is.na(Educational_level)),
  n_Income_level_missing = sum(is.na(Income_level)),
  n_URBANICITY_missing = sum(is.na(URBANICITY)),
  n_AGE_YRS_missing = sum(is.na(AGE_YRS)),
  n_SEX_OF_RESPONDENT = sum(is.na(SEX_OF_RESPONDENT)),
  n_Race_ethnicity_missing = sum(is.na(Race_ethnicity)),
  n_Marital_Status_missing =sum(is.na(Marital_Status))
  
)

colSums(is.na(Capstone_1))


```

##Further ttabulation of the missingness
```{r}
library(dplyr)
missing_summary <- Capstone_1 %>%
  summarise(across(
    everything(),
    list(
      n_missing = ~sum(is.na(.)),
      pct_missing = ~mean(is.na(.)) * 100
    ),
    .names = "{.col}_{.fn}"
  )) %>%
  tidyr::pivot_longer(everything(),
                      names_to = c("variable", ".value"),
                      names_sep = "_")

print(missing_summary)

```

##Commputing analytic sample
```{r cars}
total_n <- nrow(Capstone_1)
library(tidyr)

Capstone_Analytic_sample <-Capstone_1 %>%
  drop_na(all_of(c(
    "FIREARMS_Ownership",
    "Educational_level",
    "Income_level",
    "URBANICITY",
    "AGE_YRS",
    "SEX_OF_RESPONDENT",
    "Race_ethnicity",
    "Marital_Status"
    )))
  

```


##Loading my package
```{r}
library(dplyr)
library(gtsummary)

```


##Generating Table 1

```{r}
Capstone_Analytic_sample$FIREARMS_Ownership <-factor(Capstone_Analytic_sample$FIREARMS_Ownership,levels = c("Yes","No"))

table1 <- Capstone_Analytic_sample %>%
  select(FIREARMS_Ownership, Educational_level, Income_level, URBANICITY,
         AGE_YRS, SEX_OF_RESPONDENT, Race_ethnicity, Marital_Status) %>%
  tbl_summary(
    by = FIREARMS_Ownership,        
    statistic = list(
      all_categorical() ~ "{n} ({p}%)",
      all_continuous()  ~ "{mean} ({sd})"
    )
  ) %>%
  add_overall() %>% 
  modify_header(label = "**Variable**") %>% 
  modify_spanning_header(all_stat_cols() ~ "**Firearm Ownership**") %>%
  modify_caption("**Table 1. Descriptive Characteristics of U.S. Adult Stratified by Firearm Ownership**")

table1
```


## Exporting Table 1 to Ms WOrd
```{r}
install.packages("gt")
library(gt)

# Convert gtsummary table to gt object
gt_table <- as_gt(table1)

# Save as Word
gtsave(gt_table, "Table1_BRFSS.docx")

# Or save as HTML
gtsave(gt_table, "Table1_BRFSS.html")

R.version.string
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
